\textbf{COMPARISON OF \\ HEURISTIC ALGORITHMS \\ FOR KP01} \\[0.5cm] \rule{12cm}{0.3mm} \\[0.5cm] \small \scshape{A comparison of heuristic algorithms for solving the 0-1 Knapsack Problem} \\[0.4cm] \rule{12cm}{0.3mm} \vskip -0.3cm

### 2023-04-05 doc/paper.tex
fixed grammar and added kp01 dp pseudocode

\begin{gather}
    \text{Maximize} \sum_{i = 1}^{n} p_i x_i \\
    \text{subject to} \sum_{i = 1}^{n} w_i <= W \\
    x \in \{1, 0\}
\end{gather}

@@

\subsubsection*{Big O notation}
The time complexity of an algorithm is a function of the size of the input of that algorithm. For example, if $n$ numbers $a_1, a_2, a_3, ..., a_n$ are given, and an algorithm checks the existence of an certain element in those numbers, by checking all n elements, then the time complexity is some constant times n - the size of the input. Computer scientists would denote this using Big O notation as $O(n)$. Big O notation is a system developed by mathematicians to describe a functions asymptotic limiting factor.

@@

% time complexities graph
\begin{center}

\begin{tikzpicture}
  \begin{axis}[
    grid = major,
    clip = true,
    ticks = none,
    width=1\textwidth,
    height=0.8\textwidth,
    every axis plot/.append style={very thick},
    axis line style = ultra thick,
    clip mode=individual,
    restrict y to domain=0:10,
    restrict x to domain=0:10,
    axis x line = left,
    axis y line = left,
    domain = 0.00:10,
    xmin = 0,
    xmax = 11,
    ymin = 0,
    ymax = 11,
    xlabel = n,
    ylabel = no. of operations,
    xlabel style = {at={(axis description cs:0.5,-0.1)},anchor=south},
    ylabel style = {at={(axis description cs:-0.08,0.5)},anchor=north},
    label style = {font=\LARGE\bf},
  ]
\addplot [
    samples=100, 
    color=green,
]
{x}node[above,pos=1,style={font=\Large}]{$\mathcal{O}(n)$};
\addplot [
    samples=100, 
    color=blue,
]
{log2 x}node[above,pos=1,style={font=\Large}]{$\mathcal{O}(\log{}n)$};
\addplot [
    samples=100, 
    color=red,
]
{3*2^x}node[above,pos=1,style={font=\Large}]{$\mathcal{O}(3*2^x)$};
\addplot [
    samples=100, 
    color=black,
]
{1}node[above,pos=1,style={font=\Large}]{$\mathcal{O}(1)$};

\addplot [
  samples=100,
  color=orange, 
] gnuplot{gamma(x+1)} node[above,pos=1,style={font=\Large}]{$\mathcal{O}(n!)$};

\end{axis}
\end{tikzpicture}

\end{center}

>>>>

\vskip -0.5cm

\begin{gather}
    \text{Maximize} \sum_{i = 1}^{n} p_i x_i \\
    \text{subject to} \sum_{i = 1}^{n} w_i x_i <= W \\
    x \in \{0, 1\}
\end{gather}

@@

\subsubsection*{Big O notation}
The time complexity of an algorithm is a function of the size of the input of that algorithm. For example, if $n$ numbers $a_1, a_2, a_3, ..., a_n$ are given, and an algorithm checks the existence of an certain element in those numbers, by checking all n elements, then the time complexity is some constant times $n$ - the size of the input. Computer scientists would denote this using Big O notation as $O(n)$. Big O notation is a system developed by mathematicians and computer scientists to describe an function/algorithm's asymptotic limiting factor.

@@

% time complexities graph
\begin{center}
    \begin{tikzpicture}
      \begin{axis}[
          grid = major,
          clip = true,
          ticks = none,
          width=1\textwidth,
          height=0.8\textwidth,
          every axis plot/.append style={very thick},
          axis line style = ultra thick,
          clip mode=individual,
          restrict y to domain=0:10,
          restrict x to domain=0:10,
          axis x line = left,
          axis y line = left,
          domain = 0.00:10,
          xmin = 0,
          xmax = 11,
          ymin = 0,
          ymax = 11,
          xlabel = n,
          ylabel = Operations,
          xlabel style = {at={(axis description cs:0.5,-0.1)},anchor=south},
          ylabel style = {at={(axis description cs:-0.08,0.5)},anchor=north},
          label style = {font=\LARGE\bf},
          ]
          \addplot [
          samples=100, 
          color=green,
          ]
          {x}node[above,pos=1,style={font=\Large}]{$\mathcal{O}(n)$};
          \addplot [
          samples=100, 
          color=blue,
          ]
          {log2 x}node[above,pos=1,style={font=\Large}]{$\mathcal{O}(\log{}n)$};
          \addplot [
          samples=100, 
          color=red,
          ]
          {3*2^x}node[above,pos=1,style={font=\Large}]{$\mathcal{O}(3*2^n)$};
          \addplot [
          samples=100, 
          color=violet,
          ]
          {1}node[above,pos=1,style={font=\Large}]{$\mathcal{O}(1)$};  
          \addplot [
          samples=100,
          color=orange, 
          ] gnuplot{gamma(x+1)} node[above,pos=1,style={font=\Large}]{$\mathcal{O}(n!)$};  
      \end{axis}
    \end{tikzpicture}
\end{center}

\subsection{Solving the 0-1 Knapsack Problem}

\subsubsection*{Deterministic algorithms}
Dynamic programming is a general technique for solving optimization problems. If a problem has optimal substructure and over-lapping subproblems, then dynamic programming is applicable. In computer science 
a problem has \emph{optimal substructure} if an optimal solution can be made from optimal solutions of its sub-problems, and over-lapping subproblems is when a problem can be decomposed into sub-problems which are reused. Dynamic programming breaks down a complicate problem into smaller sub-problems in a recursive manner, while also using some memory to save the solutions of the sub-problems (either in tabular form or memoized form). This way, when we need to get a solution for a sub-problem again, we can just use our previously calculated value.

KP01 is solved using dynamic programming using a table based computation:
\begin{algorithm}
\caption{Pseudocode for solving 0-1 Knapsack with Dynamic Programming}\label{dp}
    \begin{algorithmic}
        \For {$i = 0 \text{ to } noItems $}
            \State $Table[i][0] \gets 0 $
        \EndFor
        \For {$k = 0 \text{ to } maxCapacity $}
            \State $Table[0][k] \gets 0 $
        \EndFor
        \For {$i = 0 \text{ to } noItems$}
            \For {$k = 0 \text{ to } maxCapacity $}
                \State $Table[i][k] \gets Table[i-1][k]$
                \If {$k-weights[i] >= 0$}
                    \State $Table[i][k] \gets max(Table[i-1][k-weights[i]] + profits[i], Table[i][k]) $
                \EndIf
            \EndFor
        \EndFor
        \State $\text{Print } Table[noItems][maxCapacity]$
    \end{algorithmic}
\end{algorithm}
